---
title: "Creating R Packages"
author: "Fabian Pawelczyk and Niklas Pawelzik"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open a template package:

In order to create a package, perform the following steps:

    "File", "New Project", "New Directory", "R-Package"

        Set Name, Set Sub-directory

A package template is created, go ahead and explore it!

```{r}
# insert GIF of path to Package Creation
# insert GIF or Screenshot of Package Anatomy/structure 
```

ADD COMMENT THAT WE WILL NOT PRE-LOAD PACKAGES, but will use "devtools" and "usethis"

Load your helper packages (after ensuring they are installed)

```{r}
# install.packages(c("devtools", "usethis"))
# library(devtools)
```

## Create a first function for your package

Create new function in the package you created, name it "basic_summary"

```{r}
# usethis::use_r("FUNCTION_NAME")
# FIGURE OUT: usethis::use_git("FUNCTION_NAME")
# FIGURE OUT: usethis::use_github("FUNCTION_NAME")
```

Insert basic functions into the newly created .R file

```{r}
# function that returns:
  # 1. box plot and histogram of the distribution of the input
basic_summary <- function(x){
  par(mrow=c (1, 2))
  boxplot(x)
  hist(x)
  par(mrow=c (1, 1))
  # 2. a data frame with a basic statistical description of the input
  data.frame(
    min = min(x),
    max = max(x),
    med = median(x),
    mean = mean(x)
)
  }
```

## Update your package structure

Connect our function with the rest of our package structure, use "Code", "Insert Roxygen Skeleton"

In order to take full advantage of Roxygen support, we delete the pre-existing "NAMESPACE"-file

Correctly filled out, the Rogygen comments allow you to update namespace, man and description(?)

```{r}
# delete pre-existing "NAMESPACE"-file
# use "Code", "Insert Roxygen Skeleton"
# Title of the function w/ @title
# Description of the function purpose with @description
# Documenting the function arguments/parameters with @param
# Specifying it for export with @export
# If it requires external packages, reference the package or specific package functions with @import and @importFrom
# What does the function return with @return
```

Next, you can let roxygen implement these changes into your package structure

```{r}
# devtools::document()
```

Safe this file and check whether it is correctly stored in your package´s R-folder

```{r}
# insert GIF of going to R-folder and seeing package created called basic_summary.R
```

Congrats, you created your first function within your first own R-Package!

## Check your work

Next, let´s confirm this worked as intended.

When it comes to interacting with the functions in our package, let´s establish good habits early on:

Please start by clearing objects from your environment and use the following code load them again:

```{r}
# devtools::load_all()
```

### Question: Can you explain why this approach is the best way to interact with the newly created functions in your package?

```{r}
# difficult question, possibly delete. Aims for: Makes sense for instance when we use a lot of functions in our package, ensure interactions and links between functions work as intended
```

### Question: Can you spot a difference from using load_all() for the functions in your package and creating a function in a regular R-Studio session?

```{r}
# ANSWER: The loaded functions from "load_all()" do not appear in the global environment 
```

Once you loaded all functions from your package again, we are ready to actually test them.

For this, again we establish good practices from the beginning by introducing package testing code:

```{r}
# devtools::check()
# Note: This checks only for a standardized set of formal errors
# Very useful, but we will usually want to add further tests (using "usethis::use_testthat()") that ensure our function works in the specific way we intended. Limited time, we cannot cover this part in this workshop 
```

### Question: Do you receive any errors or warnings? If so, do you understand them? Can you check and correct them?

```{r}
# GIF/screenshot of warnings?
# take care of license issue: "usethis::use_mit_license()"
# depending on whether we deleted the "hello_world" function, we might get an error message
```

Install and load your package and test it with a dataset you load!

```{r}
# devtools::install() 
# library(INSERT_PACKAGE_NAME)
# GIF/screenshot of warnings?
```

Ask RStudio for help on your function

```{r}
# ?basic_summary
# GIF/screenshot?
```

### Optitonal: Can you load your package into your Github?

```{r}
# GIF/screenshot?
```

Note that the `echo = FALSE` parameter prevents printing of the R code that generated the plot.
